Assessment of a Shen-Fourier turbulent channel flow solver for large-scale simulations

### 2016-12-06 paper.tex
Algorithm2e

\documentclass[11pt, oneside]{elsarticle}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{cancel}
\usepackage{caption}
\usepackage{subfigure}
\usepackage{algorithm}
\usepackage{algpseudocode}

\subfiglabelskip=0pt

%\newcommand{\codetitlestyle}[1]{\small\textit{#1}\hspace{0.1cm}}
\newcommand{\belowtitleskip}{2pt}%\smallskipamount}
%\newcommand{\captionposition}{t}

@@

\section{Implementation}
\label{sec:implementation}
An outline of the solution procedure used, for the numerical method described in Sec.~\ref{sec:discretizationNS}, is given in Algorithm \ref{alg:NS}.
\begin{algorithm}
	\caption{Solution procedure for Navier Stokes equations.     \label{alg:NS}}
    \begin{algorithmic}[1]
    	\State \text{Initialize} $\bm{u}(\bm{x}, 0)$ and compute 
    	$\hat{g}_{\bm{k}}(0)$ and $\hat{u}_{\bm{k}}(0)$
		\State \text{Compute LU decomposition of} $\N{H}(m,n), \D{H}(m,n) 
		\,\forall \, m, n \in K^p$
		\State \text{Compute LU decomposition of coefficient matrices in} (\ref{eq:v00}) and (\ref{eq:w00})
		\State $t=0; {\kappa}=0$
    	\While{$t<T$}
	    	\State \text{Compute nonlinear convection} $\bm{\hat{\mathcal{H}}}^{\kappa}$
	    	\State \text{Compute right hand sides of Eqs.~}(\ref{eq:ufin}), 
	    	(\ref{eq:gfin})
		    \State \text{Solve Eq.}(\ref{eq:ufin}) for 
		    $\hat{u}^{{\kappa}+1}_{\bm{k}} \, \forall \,\bm{k} \in \N{K}_N$
		    \State \text{Solve Eqs.}(\ref{eq:gfin}) for 
		    $\hat{g}^{{\kappa}+1}_{\bm{k}} \, \forall\, \bm{k} \in \D{K}_N$
		    \State \text{Solve Eqs.}(\ref{eq:f3}, \ref{eq:g3}) for  
		    $\hat{v}_{\bm{k}}, \hat{w}_{\bm{k}}\, \forall\, \bm{k} 
		    \in \D{K}_N^0$
		    \State \text{Solve Eqs.}(\ref{eq:v00}, \ref{eq:w00}) for 
		    $\hat{v}_0^{{\kappa}+1}, \hat{w}_0^{{\kappa}+1}\, 
		    \forall \, l \in \D{K}^x$
		    \State ${\kappa} \gets {\kappa} + 1$
		    \State $t \gets t+\triangle t$
		    \State \text{Update to new time step}
	    \EndWhile

    \end{algorithmic}
\end{algorithm}
The major computational cost comes from  computing the nonlinear convection 
and solving for Eqs.~(\ref{eq:ufin}, \ref{eq:gfin}). The nonlinear 
convection 
term, $\bm{\hat{\mathcal{H}}}$, is computed using a pseudo-spectral method with 
the 3/2-rule for dealiasing in the periodic directions, see \cite{canuto1988}. The 
transforms are carried out using fast Fourier and fast Chebyshev transforms 
that are all of order $\mathcal{O}(N_x\log N_x)$ per one-dimensional transform of size 
$N_x$. The Shen transforms make use of the Chebyshev transforms that are using 
discrete cosine transforms. The forward scalar products and transforms are implemented as shown 
in Algorithm \ref{alg:fst} of the Appendix. A complete transform is obtained through a scalar 
product followed by solving the linear system involving the mass 
matrix, just like shown in Eq. (\ref{eq:fstB}) for space $\N{V}_N$ and similar 
for $\D{V}_N$ and $V_N$. Since the mass matrices are diagonal, tridiagonal and 
pentadiagonal for $V_N, \D{V}_N$ and $\N{V}_N$ respectively, solving the linear 
system is $\mathcal{O}(N_x)$ and the complete transform still scales like $\mathcal{O}(N_x\log N_x)$.

@@

\begin{algorithm}
	\caption{LU decomposition of biharmonic operator $\N{H}$. Current algorithm 
	is using dense storage of $\N{U}_{kj}$, but no more than the 3 items per row
	need 
	to be stored after the row is no longer needed in the for-loop below (as the loop is traversed we do not need $\N{U}_{m, j}$ for $m<k-4$). \label{alg:lu_biharmonic}}
	\begin{algorithmic}[1]
		\Procedure{LUbiharmonic}{}
			\State \textbf{Input:} $\{\N{H}_{kj}\}_{k, j = 0}^{N}$
			\State Integer $M_e \gets \lfloor N/2 \rfloor$
			\State Integer $M_o \gets \lfloor (N-1)/2 \rfloor$  
	        \State $\{\hat{U}_{0,2j}\}_{j=0}^{M_e} \gets \{\hat{H}_{0,2j}\}_{j=0}^{M_e}$ 
	        \State $\{\hat{U}_{1,2j+1}\}_{j=0}^{M_o} \gets \{\hat{H}_{1,2j+1}\}_{j=0}^{M_o}$ 
	        \State $\N{L}_{20} \gets \N{H}_{20}/\N{U}_{00}$
	        \State $\N{L}_{31} \gets \N{H}_{31}/\N{U}_{11}$
	        \State $\{\hat{U}_{2,2j}\}_{j=1}^{M_e} \gets \{\N{H}_{2,2j}\}_{j=1}^{M_e} - \N{L}_{20} \{\N{U}_{0,2j}\}_{j=1}^{M_e}$
	        \State $\{\hat{U}_{3,2j+1}\}_{j=1}^{M_o} \gets \{\N{H}_{3,2j+1}\}_{j=1}^{M_o} - \N{L}_{31} \{\N{U}_{1,2j+1}\}_{j=1}^{M_o}$

	        %\State $\N{U}_{2j} = \N{H}_{2j} - \N{L}_{20} \N{U}_{0j} \quad 
	        %\forall \, j \in 2, 4, \ldots, N$
	        %\State $\N{U}_{3j} = \N{H}_{3j} - \N{L}_{31} \N{U}_{1j} \quad 
	        %\forall \, j \in 
	        %3, 5, \ldots, N$
	        \For{$k = 4, 5, \ldots, N$}
	            \State Integer $M \gets \lfloor \frac{N-k}{2}\rfloor$
		        \State $\N{L}_{k,k-4} \gets \N{H}_{k, k-4} / \N{U}_{k-4, k-4}$
		        \State $\{\N{U}_{k,k+2j}\}_{j=-1}^{M} \gets \{\N{H}_{k,k+2j}\}_{j=-1}^{M} - \N{L}_{k, k-4} 
		        \{\N{U}_{k-4, k+2j}\}_{j=-1}^{M}$
		        %\State $\N{U}_{k,j} = \N{H}_{k,j} - \N{L}_{k, k-4}  \N{U}_{k-4, j} \quad \forall j = k-2, k, \ldots, N$
	            \State $\N{L}_{k,k-2} \gets {\N{U}}_{k,k-2}/{\N{U}}_{k-2, k-2}$
	            \State $\{\N{U}_{k,k+2j}\}_{j=0}^{M} \gets \{{\N{U}}_{k,k+2j}\}_{j=0}^{M} - \N{L}_{k, k-2} \{\N{U}_{k-2,k+2j}\}_{j=0}^{M}$
		        %\State $\N{U}_{kj} = {\N{U}}_{kj} - \N{L}_{k, k-2} \N{U}_{k-2, j} \quad \forall j = k, k+2, \ldots, N$
	        \EndFor
	        \State \Return $\{\N{L}_{kj}\}_{k, j = 0}^{N}$, $\{\N{U}_{kj}\}_{k, 
	        j = 
	        0}^{N}$
		\EndProcedure	
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}
	\caption{Recursive formula to compute $\{a_k\}_{k=0}^{N_x-4}, 
	\{b_k\}_{k=0}^{N_x-4}$ in Eq. (\ref{eq:ab}). Note that the lower triangular 
	matrix $\{L_{kj}\}_{k,j=0}^{N_x-4}$ contains only two nonzero diagonals and 
	the parameters $p_k$ and $r_k$ are given in Eqs. (\ref{eq:pk}) and 
	(\ref{eq:rk}) respectively. \label{alg:ab}}
	\begin{algorithmic}[1]
		\Procedure{ReduceAB}{}
		\State \textbf{Input:} $ \{\N{L}_{kj}\}_{k,j=0}^{N_x-4}, \xi_0$
		\For{$k=0,1$}
		     \State $a_k \gets p_k$
		     \State $b_k \gets r_k$
		\EndFor
		\For{$k=2,3$}
		    \State $a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2}$
		    \State $b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}$
		\EndFor
		\For{$k=4, 5, \ldots, N_x-4$}
     		\State $a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2} - \N{L}_{k, k-4} 
\,a_{k-4}$
	    	\State $b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}- \N{L}_{k, k-4} 
\,b_{k-4}$
		\EndFor

		\State \Return $\{a_k\}_{k=0}^{N_x-4}, \{b_k\}_{k=0}^{N_x-4}$
        \EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
	\caption{Solve biharmonic Eq. (\ref{eq:ufin}) with pre-computed $\{\N{U}_{kj}\}_{k,j=0}^{N}, 
		\{\N{L}_{kj}\}_{k,j=0}^{N}$ 
	matrices, as well as the $\{ a_k\}_{k=0}^N$ and $\{ b_k\}_{k=0}^N$ vectors. The parameters $q_j$ and $s_j$ are given in Eqs. (\ref{eq:pk}) and 
	(\ref{eq:rk}) respectively. 	\label{alg:SolveBiharmonic}}
	\begin{algorithmic}[1]
		\Procedure{LUSolveBiharmonic}{}
		  \State \textbf{Input:} $\{\N{U}_{kj}\}_{k,j=0}^{N}, 
		  \{\N{L}_{kj}\}_{k,j=0}^{N}, 
		  \{f_k\}_{k=0}^{N}, \{a_k\}_{k=0}^{N}, \{b_k\}_{k=0}^{N}, \xi_0$ 
		  \State // Solve $\N{L}_{kj}y_j=f_k$ by forward elimination 
		  \State $y_0 \gets f_0$
		  \State $y_1 \gets f_1$
		  \State $y_2 \gets f_2 - \N{L}_{20}y_{0}$
		  \State $y_3 \gets f_3 - \N{L}_{31}y_{1}$
		  \For{$k = 4$ \textbf{to} $N$}
		  \State $y_k \gets f_k - \N{L}_{k,k-2}y_{k-2} - \N{L}_{k, k-4}y_{k-4}$
		  \EndFor
		  \State // Solve $\N{U}_{kj}u_j=y_k$ with back 
		  substitution                    
		  \State $u_N \gets y_N / \N{U}_{NN}$
	      \State $u_{N-1} \gets y_{N-1} / \N{U}_{N-1,N-1}$
	      \State $u_{N-2} \gets \left(y_{N-2} - \N{U}_{N-2, 
	      N}y_N\right)/\N{U}_{N-2,N-2}$
	      \State $u_{N-3} \gets \left(y_{N-3} - \N{U}_{N-3, 
	      N-1}y_{N-1}\right)/\N{U}_{N-3,N-3}$

	      \State Float $q^o = q^e = s^o = s^e = 0.0$
		  \For{$k = N-4$ \textbf{to} $0$ \textbf{step} $-1$}
		  \State $u_k \gets y_k - \N{U}_{k, k+2} u_{k+2} - \N{U}_{k, k+4} 
		  u_{k+4} $ 
		  \If{$k < N-5$}
		  \State $j \gets k+6$
		  \If{$k$ is odd}
		    \State $q^o \gets q^o + u_{j}q_{j}$
		    \State $s^o \gets s^o + u_{j}s_{j}$
		    \State $u_k \gets u_k - \xi_0q^o a_k - \xi_0s^o b_k $	    
		  \Else
		    \State $q^e \gets q^e + u_{j}q_{j}$
		    \State $s^e \gets s^e + u_{j}s_{j}$	  
		    \State $u_k \gets u_k - \xi_0q^e a_k - \xi_0s^e b_k $
		  \EndIf
		  \EndIf
		  \State $u_k \gets  u_k / \N{U}_{kk}$
		  \EndFor
		  \State \Return $\{u_k\}_{k=0}^{N}$

		\EndProcedure
	\end{algorithmic}
\end{algorithm}

@@

\bibliography{bib.bib}

@@

The algorithms for the fast scalar products and transforms, for the three spaces $V_N, \D{V}_N, \N{V}_N$, are given in Alg. ~\ref{alg:fst}. The algorithms for the inverse transforms are given in Alg.~\ref{alg:ifst}. The algorithms are using the discrete cosine transforms (dct) of type 1, 2 or 3 depending on the choice of discretization points and direction.
\begin{algorithm}
	\caption{Forward scalar products and transforms for all spaces $V_N, \D{V}_N, \N{V}_N$. 
		Here "dct" is the discrete cosine transform from SciPy.}
		\label{alg:fst}
		\begin{algorithmic}[1]
			\Function{forwardScalar}{$\{f_j\}_{j=0}^{N}$, points, space}
			\If{ points = "Gauss-Lobatto"}
			\State $\{ w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, \text{type=2}, 
			\text{axis=0})\frac{\pi}{2 N}$
			\Else{ points = "Gauss-Chebyshev"}
			\State $\{ w_k\}_{k=0}^{N} \gets \text{dct}(f, \text{type=1}, 
			\text{axis=0}) \frac{\pi}{2 (N-1)}$
			\EndIf
			\State $\{y\}_{k=0}^N \gets 0$
			\If{ space = $\D{V}_N$}
			\State $\{y_k\}_{k=0}^{N-2} \gets \{w_k\}_{k=0}^{N-2} - \{w_k\}_{k=2}^{N}$
			\ElsIf{ space = $\N{V}_N$}	                
			\For{$k=0$ to $N-4$}
			\State ${y}_k \gets w_k - \frac{2(k+2)}{k+3} w_{k+2} + 
			\frac{k+1}{k+3}w_{k+4}$
			\EndFor
			\ElsIf{space = $V_N$}
			\State $\{y_k\}_{k=0}^N \gets \frac{2}{\pi} \{w_k\}_{k=0}^N $
			\State $y_0 \gets y_0 / 2$
			\If{ points = "Gauss-Lobatto"}
			\State $y_N \gets y_N / 2$	     	
			\EndIf
			\EndIf
			\State \Return $\{y_k\}_{k=0}^{N}$    
			\EndFunction
			\Function{forwardTransform}{$\{f_j\}_{j=0}^{N}$, points, space}
			\State $\{y_k\}_{k=0}^{N} \gets$ forwardScalar($\{f_j\}_{j=0}^{N}$, points, space)
			\If{space = $\D{V}_N$}
			\State $B \gets \D{B}$
			\ElsIf{space = $\N{V}_N$}
			\State $B \gets \N{B}$
			\ElsIf{space = $V_N$}
			\State $B \gets B$
			\EndIf
			\State $\bm{y} \gets B^{-1} \bm{y}$
			\State \Return $\{y_k\}_{k=0}^{N}$
			\EndFunction

		\end{algorithmic}
\end{algorithm}

\begin{algorithm}
	\caption{Inverse transforms for all spaces $V_N, \D{V}_N, \N{V}_N$. 
		Here "dct" is the discrete cosine transform from SciPy.}
	\label{alg:ifst}
	\begin{algorithmic}[1]
		\Function{inverseChebTransform}{$\{f\}_{j=0}^N$, points}
		\If{points = "Gauss-Chebyshev"}
		    \State $\{w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, type=3, axis=0)$
		    \State $\{w_k\}_{k=0}^{N} \gets \frac{1}{2}\{w_k\}_{k=0}^{N} + \frac{1}{2}f_0$

        \ElsIf{ points = "Gauss-Lobatto"}
	        \State $\{w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, type=1, axis=0)$
	        \State $\{w_k\}_{k=0}^{N} \gets \frac{1}{2}\{w_k\}_{k=0}^{N} + \frac{1}{2}f_0$
	        \For{$k=0$ \textbf{to} $N$}
	         \State $w_k \gets \frac{(-1)^k}{2} f_N$
	        \EndFor
	    \EndIf
	    \State \Return{$\{w_k\}_{k=0}^{N}$}

	    \EndFunction

		\Function{inverseShenTransform}{$\{f_j\}_{j=0}^{N}$, points, space}

		\If{ space = $\D{V}_N$}
		\State $\{w_k\}_{k=0}^{N-2} \gets \{f_k\}_{k=0}^{N-2}$
		\State $\{w_k\}_{k=2}^{N} \gets \{w_k\}_{k=2}^{N} - \{f_k\}_{k=0}^{N-2} $
        \ElsIf{ space = $\N{V}_N$}	                
		\State $\{w_k\}_{k=0}^{N-4} \gets \{f_k\}_{k=0}^{N-4}$
		\For{$k=0$ \textbf{to} $N-4$}
		\State $w_{k+2} \gets w_{k+2} - \frac{2(k+2)}{k+3} f_k$
		\EndFor
		\For{$k=0$ \textbf{to} $N-4$}
		\State $w_{k+4} \gets w_{k+4} + \frac{k+1}{k+3} f_k$
		\EndFor
        \EndIf
        \State $\{{w}_k\}_{k=0}^{N}$ = inverseChebTransform($\{w_k\}_{k=0}^N$, points)
		\State \Return $\{{w}_k\}_{k=0}^{N}$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

>>>>

\documentclass[preprint]{elsarticle}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{cancel}
\usepackage{caption}
\usepackage{subfigure}
\usepackage[ruled,vlined]{algorithm2e}
%\usepackage{algpseudocode}

@@

\section{Implementation}
\label{sec:implementation}
An outline of the solution procedure used, for the numerical method described 
in Sec.~\ref{sec:discretizationNS}, is given in Algorithm \ref{alg:NS}. Note 
that the Adams-Bashforth projection of convection reverts to a first order 
forward Euler on the first time step. For simulations of turbulent flows this 
is insignificant, but for accuracy tests, like the Orr-Sommerfeld instability, 
a proper initialization would need to set both the two first time steps instead.
\begin{algorithm}
	\caption{Solution procedure for Navier Stokes equations.    
	 \label{alg:NS}}
 	\DontPrintSemicolon	
    	Initialize $\bm{u}(\bm{x}, 0)$ and compute 
    	$\hat{g}_{\bm{k}}(0)$ and $\hat{\bm{u}}_{\bm{k}}(0)$ \;
		Compute LU decomposition of $\N{H}(m,n), \D{H}(m,n) 
		\,\forall \, m, n \in K^p$ \;
		Compute LU decomposition of coefficient matrices in (\ref{eq:v00}) and 
		(\ref{eq:w00}) \;
		$t \gets 0$\; 
		${\kappa} \gets 0$ \;
    	Compute nonlinear convection 
    	$\bm{\hat{\mathcal{H}}}^{\kappa-1}(\hat{\bm{u}}_{\bm{k}}(0), 
    	\hat{g}_{\bm{k}}(0))$ \;
    	\While{$t<T$}{
	    	Compute nonlinear convection $\bm{\hat{\mathcal{H}}}^{\kappa}$ \;
	    	{Compute right hand sides of Eqs.~}(\ref{eq:ufin}), 
	    	(\ref{eq:gfin}) \;
		    {Solve Eq.}(\ref{eq:ufin}) for 
		    $\hat{u}^{{\kappa}+1}_{\bm{k}} \, \forall \,\bm{k} \in \N{K}_N$ \;
		    {Solve Eqs.}(\ref{eq:gfin}) for 
		    $\hat{g}^{{\kappa}+1}_{\bm{k}} \, \forall\, \bm{k} \in \D{K}_N$ \;
		    {Solve Eqs.}(\ref{eq:f3}, \ref{eq:g3}) for  
		    $\hat{v}_{\bm{k}}, \hat{w}_{\bm{k}}\, \forall\, \bm{k} 
		    \in \D{K}_N^0$ \;
		    {Solve Eqs.}(\ref{eq:v00}, \ref{eq:w00}) for 
		    $\hat{v}_0^{{\kappa}+1}, \hat{w}_0^{{\kappa}+1}\, 
		    \forall \, l \in \D{K}^x$ \;
		    ${\kappa} \gets {\kappa} + 1$ \;
		    $t \gets t+\triangle t$ \;
		    {Update to new time step} \;
	    }
\end{algorithm}

The major computational cost in Alg. \ref{alg:NS} comes from  computing the 
nonlinear convection and solving for Eqs.~(\ref{eq:ufin}, \ref{eq:gfin}). The 
nonlinear convection term, $\bm{\hat{\mathcal{H}}}$, is computed using a 
pseudo-spectral method with the 3/2-rule for dealiasing in the periodic 
directions, see \cite{canuto1988}. The 
transforms are carried out using fast Fourier and fast Chebyshev transforms 
that are all of order $\mathcal{O}(N_x\log N_x)$ per one-dimensional transform of size 
$N_x$. The Shen transforms make use of the Chebyshev transforms that are using 
discrete cosine transforms. The forward scalar products and transforms are implemented as shown 
in Algorithm \ref{alg:fst} of the Appendix. A complete transform is obtained through a scalar 
product followed by solving the linear system involving the mass 
matrix, just like shown in Eq. (\ref{eq:fstB}) for space $\N{V}_N$ and similar 
for $\D{V}_N$ and $V_N$. Since the mass matrices are diagonal, tridiagonal and 
pentadiagonal for $V_N, \D{V}_N$ and $\N{V}_N$ respectively, solving the linear 
system is $\mathcal{O}(N_x)$ and the complete transform still scales like $\mathcal{O}(N_x\log N_x)$.

@@

\SetKwProg{Fn}{Function}{}{}
\SetKwFunction{reduceAB}{ReduceAB}
\begin{algorithm}
	\DontPrintSemicolon
    \caption{Recursive formula to compute $\{a_k\}_{k=0}^{N_x-4}, 
	\{b_k\}_{k=0}^{N_x-4}$ in Eq. (\ref{eq:ab}). Note that the lower triangular 
	matrix $\{L_{kj}\}_{k,j=0}^{N_x-4}$ contains only two nonzero diagonals and 
	the parameters $p_k$ and $r_k$ are given in Eqs. (\ref{eq:pk}) and 
	(\ref{eq:rk}) respectively. \label{alg:ab}}
		\Fn{\reduceAB{$ \{\N{L}_{kj}\}_{k,j=0}^{N_x-4}, \,\xi_0$}}{
		\For{$k=0,1$}{
		     $a_k \gets p_k$ \;
		     $b_k \gets r_k$ \;
		}
		\For{$k=2,3$}{
		    $a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2}$ \;
		    $b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}$ \;
		}
		\For{$k=4, 5, \ldots, N_x-4$}{
     		$a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2} - \N{L}_{k, k-4} \,a_{k-4}$ 
     		\;
	    	$b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}- \N{L}_{k, k-4} \,b_{k-4}$ 
	    	\;
		}
		\Return $\{a_k\}_{k=0}^{N_x-4}, \{b_k\}_{k=0}^{N_x-4}$ \;
    }
\end{algorithm}

\SetKwFunction{LUSolveBiharmonic}{LUSolveBiharmonic}
\begin{algorithm}
    \DontPrintSemicolon
	\caption{Solve biharmonic Eq. (\ref{eq:ufin}) with pre-computed 
	$\{\N{U}_{kj}\}_{k,j=0}^{N}, 
		\{\N{L}_{kj}\}_{k,j=0}^{N}$ 
	matrices, as well as the $\{ a_k\}_{k=0}^N$ and $\{ b_k\}_{k=0}^N$ vectors. 
	The parameters $q_j$ and $s_j$ are given in Eqs. (\ref{eq:pk}) and 
	(\ref{eq:rk}) respectively. 	\label{alg:SolveBiharmonic}}
		\Fn{\LUSolveBiharmonic{$\{\N{U}_{kj}\}_{k,j=0}^{N}, 
                \{\N{L}_{kj}\}_{k,j=0}^{N}, 
                \{f_k\}_{k=0}^{N}, \{a_k\}_{k=0}^{N}, \{b_k\}_{k=0}^{N}, 
                \xi_0$}}{
          \SetKwInOut{Output}{output}
          \Output{$\{ \hat{u}_k\}_{k=0}^N$}
		  \tcp*[l]{Solve $\N{L}_{kj}y_j=f_k$ by forward elimination}
		  $y_0 \gets f_0$ \;
		  $y_1 \gets f_1$ \;
		  $y_2 \gets f_2 - \N{L}_{20}y_{0}$ \;
		  $y_3 \gets f_3 - \N{L}_{31}y_{1}$ \;
		  \For{$k = 4$ \textbf{to} $N$}{
		  $y_k \gets f_k - \N{L}_{k,k-2}y_{k-2} - \N{L}_{k, k-4}y_{k-4}$ \;
		  }
		  \tcp*[l]{Solve $\N{U}_{kj}u_j=y_k$ with back 
              substitution}                    
		  $\hat{u}_N \gets y_N / \N{U}_{NN}$ \;
	      $\hat{u}_{N-1} \gets y_{N-1} / \N{U}_{N-1,N-1}$ \;
	      $\hat{u}_{N-2} \gets \left(y_{N-2} - \N{U}_{N-2, 
	      N}y_N\right)/\N{U}_{N-2,N-2}$ \;
	      $\hat{u}_{N-3} \gets \left(y_{N-3} - \N{U}_{N-3, 
	      N-1}y_{N-1}\right)/\N{U}_{N-3,N-3}$ \;

	      float $q^o = q^e = s^o = s^e = 0.0$ \;
		  \For{$k = N-4$ \textbf{to} $0$ \textbf{step} $-1$}{
		  $\hat{u}_k \gets y_k - \N{U}_{k, k+2} \hat{u}_{k+2} - \N{U}_{k, k+4} 
		  \hat{u}_{k+4} $ \;
		  \If{$k < N-5$}{
		  $j \gets k+6$ \;
		  \If{$k$ is odd}{
		    $q^o \gets q^o + \hat{u}_{j}q_{j}$ \;
		    $s^o \gets s^o + \hat{u}_{j}s_{j}$ \;
		    $\hat{u}_k \gets \hat{u}_k - \xi_0q^o a_k - \xi_0s^o b_k $ \;
        }
		  \Else{
		    $q^e \gets q^e + \hat{u}_{j}q_{j}$ \;
		    $s^e \gets s^e + \hat{u}_{j}s_{j}$ \;	  
		    $\hat{u}_k \gets \hat{u}_k - \xi_0q^e a_k - \xi_0s^e b_k $ \;
		}
		}
		  $\hat{u}_k \gets  \hat{u}_k / \N{U}_{kk}$ \;
		  }
		  \Return $\{\hat{u}_k\}_{k=0}^{N}$ \;  
		}
\end{algorithm}

@@

\bibliographystyle{model1-num-names}
\bibliography{bib.bib}

@@

The algorithms for the fast scalar products and transforms, for the three spaces $V_N, \D{V}_N, \N{V}_N$, are given in Alg. ~\ref{alg:fst}. The algorithms for the inverse transforms are given in Alg.~\ref{alg:ifst}. The algorithms are using the discrete cosine transforms (dct) of type 1, 2 or 3 depending on the choice of discretization points and direction.
\SetKwFunction{forwardScalar}{forwardScalar}
\begin{algorithm}
    \DontPrintSemicolon
	\caption{Forward scalar products and transforms for all spaces $V_N, 
	\D{V}_N, \N{V}_N$. 
		Here "dct" is the discrete cosine transform from SciPy. \label{alg:fst}}
			\Fn{\forwardScalar{$\{f_j\}_{j=0}^{N}$, points, space}}{
			\If{ points = "Gauss-Lobatto"}{
			$\{ w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, \text{type=2}, 
			\text{axis=0})\frac{\pi}{2 N}$ \;
            }
			\ElseIf{ points = "Gauss-Chebyshev"}{
			$\{ w_k\}_{k=0}^{N} \gets \text{dct}(f, \text{type=1}, 
			\text{axis=0}) \frac{\pi}{2 (N-1)}$ \;
		    }
			$\{y\}_{k=0}^N \gets 0$ \;
			\If{ space = $\D{V}_N$}{
			$\{y_k\}_{k=0}^{N-2} \gets \{w_k\}_{k=0}^{N-2} - \{w_k\}_{k=2}^{N}$ 
			\;
            }
			\ElseIf{ space = $\N{V}_N$}{
			\For{$k=0$ to $N-4$}{
			${y}_k \gets w_k - \frac{2(k+2)}{k+3} w_{k+2} + 
			\frac{k+1}{k+3}w_{k+4}$ \;
		    }
            }
			\ElseIf{space = $V_N$}{
			$\{y_k\}_{k=0}^N \gets \frac{2}{\pi} \{w_k\}_{k=0}^N $ \;
			$y_0 \gets y_0 / 2$ \;
            }
			\If{ points = "Gauss-Lobatto"}{
			$y_N \gets y_N / 2$ \;   	
		    }
			\Return $\{y_k\}_{k=0}^{N}$    \;
        }
%			\EndFunction
%			\Function{forwardTransform}{$\{f_j\}_{j=0}^{N}$, points, space}
%			\State $\{y_k\}_{k=0}^{N} \gets$ forwardScalar($\{f_j\}_{j=0}^{N}$, 
%points, space)
%			\If{space = $\D{V}_N$}
%			\State $B \gets \D{B}$
%			\ElsIf{space = $\N{V}_N$}
%			\State $B \gets \N{B}$
%			\ElsIf{space = $V_N$}
%			\State $B \gets B$
%			\EndIf
%			\State $\bm{y} \gets B^{-1} \bm{y}$
%			\State \Return $\{y_k\}_{k=0}^{N}$
%		}
\end{algorithm}
%
%\begin{algorithm}
%	\caption{Inverse transforms for all spaces $V_N, \D{V}_N, \N{V}_N$. 
%		Here "dct" is the discrete cosine transform from SciPy.}
%	\label{alg:ifst}
%	\begin{algorithmic}[1]
%		\Function{inverseChebTransform}{$\{f\}_{j=0}^N$, points}
%		\If{points = "Gauss-Chebyshev"}
%		    \State $\{w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, type=3, axis=0)$
%		    \State $\{w_k\}_{k=0}^{N} \gets \frac{1}{2}\{w_k\}_{k=0}^{N} + \frac{1}{2}f_0$
%		    
%        \ElsIf{ points = "Gauss-Lobatto"}
%	        \State $\{w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, type=1, axis=0)$
%	        \State $\{w_k\}_{k=0}^{N} \gets \frac{1}{2}\{w_k\}_{k=0}^{N} + \frac{1}{2}f_0$
%	        \For{$k=0$ \textbf{to} $N$}
%	         \State $w_k \gets \frac{(-1)^k}{2} f_N$
%	        \EndFor
%	    \EndIf
%	    \State \Return{$\{w_k\}_{k=0}^{N}$}
%	    
%	    \EndFunction
%		
%		\Function{inverseShenTransform}{$\{f_j\}_{j=0}^{N}$, points, space}
%		
%		\If{ space = $\D{V}_N$}
%		\State $\{w_k\}_{k=0}^{N-2} \gets \{f_k\}_{k=0}^{N-2}$
%		\State $\{w_k\}_{k=2}^{N} \gets \{w_k\}_{k=2}^{N} - \{f_k\}_{k=0}^{N-2} $
%        \ElsIf{ space = $\N{V}_N$}	                
%		\State $\{w_k\}_{k=0}^{N-4} \gets \{f_k\}_{k=0}^{N-4}$
%		\For{$k=0$ \textbf{to} $N-4$}
%		\State $w_{k+2} \gets w_{k+2} - \frac{2(k+2)}{k+3} f_k$
%		\EndFor
%		\For{$k=0$ \textbf{to} $N-4$}
%		\State $w_{k+4} \gets w_{k+4} + \frac{k+1}{k+3} f_k$
%		\EndFor
%        \EndIf
%        \State $\{{w}_k\}_{k=0}^{N}$ = inverseChebTransform($\{w_k\}_{k=0}^N$, points)
%		\State \Return $\{{w}_k\}_{k=0}^{N}$
%		\EndFunction
%	\end{algorithmic}
%\end{algorithm}
